% Document class and font size
\documentclass[a4paper,9pt]{extarticle}

% Packages
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{fancyhdr}

% configure page with geometry
\geometry{letterpaper, margin=0.75in}

% Formatting
% Removes item separation
\setlist{noitemsep}
% Section title format
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}[\titlerule]
% Section title spacing
\titlespacing*{\section}{0pt}{\baselineskip}{\baselineskip}

% Begin document
\begin{document}

% Disable page numbers
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhead{}
\fancyhead[L]{\textit{Calvin Sprouse}}
\fancyhead[R]{\textit{December 2023}}
% Remove header from the first page
\thispagestyle{empty}

% Header
\begin{flushleft}
% Name
\textbf{\LARGE Calvin Sprouse}\\[2pt]
B.S. Physics Undergraduate, Central Washington University (CWU), Ellensburg, Washington.\\
% Contact info
\href{mailto:sprousecal@gmail.com}{sprousecal@gmail.com}
 | \href{tel:4252602368}{(425) 260-2368}
 | \href{https://www.linkedin.com/in/calvin-sprouse}{linkedin.com/in/calvin-sprouse}
%  | \href{https://JohnDoe/en/staff/single/59}{Organization page}
\end{flushleft}


% \section*{RESEARCH INTERESTS}
% \noindent
% Theoretical Physics, Quantum Gravity, High Energy Physics, Astroparticle Physics, Particle Physics


\section*{EDUCATION}
\noindent
\textbf{Central Washington University}, Ellensburg, Washington \hfill September 2020 | June 2024\\
B.S. Physics \hfill Cumulative GPA: 3.73/4.00 \\
Minors in Astronomy and Mathematics


\section*{RESEARCH EXPERIENCE}
\noindent\textbf{Development of a Magneto Caloric Testing Apparatus} \hfill CWU\\
\textit{Undergraduate Student Researcher} \hfill September 2023 | Present
\begin{itemize}
    \item Developed software in Python for an Arduino controlled measurement device.
    \item Research sampling material for device calibration.
    \item Wrote funding proposals based on designs and future testing requirements.
\end{itemize}

\noindent\textbf{Computational Models of Neuronal Cytoskeleton} \hfill CWU | NSF RUI\\
\textit{Undergraduate Student Researcher} \hfill June 2022 | Present
\begin{itemize}
    \item Develop computational models on the mechanics of the neuronal Cytoskeleton in MatLab.
    \item Developed data visualizations in Python.
    \item Presented work at six research conferences including the Biophysical Society.
    \item Made posters and presentations for conferences.
    \item Wrote user guides and kept a lab notebook documenting procedures.
    \item Work supported by NSF RUI award 1915477.
\end{itemize}

\noindent\textbf{Calculating the Energy of IceCube Neutrino 190331A}\hfill University of Wisconsin-Madison | NSF REU\\
\textit{Astrophysics REU Student} \hfill May 2023 | August 2023
\begin{itemize}
    \item Developed Monte Carlo simulations for calculating neutrino energies.
    \item Worked with IceCube data.
    \item Learned to work within IceCube data analysis protocols.
    \item Learned to work with large research teams and follow data handling procedures.
    \item Presenting work at the American Astronomical Society.
\end{itemize}

\noindent\textbf{Characterization of CWU 0.6m Telescope Precision} \hfill CWU\\
\textit{Undergraduate Student Researcher} \hfill March 2022 | June 2023
\begin{itemize}
    \item Created and executed methods to reliably measure variable stars.
    \item Wrote Python scripts to automatically control a research telescope.
    \item Wrote Python scripts to preform data analysis on hundreds of observations.
    \item Presented this research at two research conferences.
\end{itemize}

\noindent\textbf{Construction of a new Spectrograph for the CWU 0.6m Research Telescope} \hfill CWU\\
\textit{Undergraduate Student Researcher} \hfill September 2022 | June 2023
\begin{itemize}
    \item Programmed data reduction notebooks in Python for an under construction spectrograph.
    \item Researched data analysis methods for using spectrograph data.
\end{itemize}


\section*{TALKS AND PRESENTATIONS}
\noindent\textbf{American Astronomical Society 243rd Meeting Poster}\\
Sprouse et.\ al. (2024). The Highest-Energy Astrophysical Muon-Neutrino: 190331A. American Astronomical Society 243rd Meeting.

\noindent\textbf{Murdock College Science Research Conference 2023 Poster}\\
Sprouse et.\ al. (2023). Characterizing the magnitude discrimination limit of the CWU 0.6m Research Telescope. Murdock College Science Research Conference 2023.

\noindent\textbf{Washington American Association of Physics Teachers Meeting 2023 Poster}\\
Title

\noindent\textbf{CWU Symposium for Undergraduate Research and Creative Expression 2023 Poster}\\
Title

\noindent\textbf{CWU Symposium for Undergraduate Research and Creative Expression 2023 Presentation}\\
Title

\noindent\textbf{Biophysical Society Meeting 2023 Poster}\\
Craig et.\ al. (2023). Self-organization of the microtubule cytoskeleton in developing axons. Biophysical Journal Abstract, \href{www.doi.org/10.1016/j.bpj.2022.11.1506}{doi.org/10.1016/j.bpj.2022.11.1506}

\noindent\textbf{Murdock College Science Research Conference 2022 Poster}\\
Title

\noindent\textbf{Physics Congress 2022 Poster}\\
Title


\section*{TEACHING EXPERIENCE}
\noindent\textbf{CWU Department of Physics} \hfill Ellensburg, Washington\\
\textit{Learning Assistant (LA) for PHYS 181 Introductory Physics with Lab} \hfill September 2023 | Present
\begin{itemize}
    \item Worked in the classroom to assist with labs.
    \item Graded student assignments and provided feedback.
    \item Provided explanations to students in class on lecture material.
    \item Read and discussed physics teaching literature as part of an accompanying class.
\end{itemize}

\noindent\textbf{CWU Learning Commons Math Center} \hfill Ellensburg, Washington\\
\textit{Lead Tutor} \hfill September 2022 | Present
\begin{itemize}
    \item Work with students by drop-in or appointment on math issues in any STEM course.
    \item Reinforce study habits with a focus on math concepts.
    \item Develop and give trainings to fellow tutors according to CRLA standards.
    \item Collect and analyze data on the center to increase tutor effectiveness and request funding.
\end{itemize}

\noindent\textbf{CWU Learning Commons Peer Assisted Learning Center} \hfill Ellensburg, Washington\\
\textit{Physics Tutor} \hfill September 2022 | Present
\begin{itemize}
    \item Provide individual and group help to students in any physics course.
    \item Reinforce study habits with a focus on physics concepts.
    \item Create session plans targeted to introductory physics students.
    \item Regularly interact with professors of introductory classes to advertise and discuss upcoming coursework.
\end{itemize}

\noindent\textbf{CWU Department of Physics} \hfill Ellensburg, Washington\\
\textit{Peer Teacher for PHYS 303 Observational Astronomy} \hfill March 2023 | May 2023
\begin{itemize}
    \item Developed astronomy focused lesson plans for a quarter long course.
    \item Taught lessons using a planetarium.
    \item Taught students how to identify constellations and use celestial coordinate systems using only their eyes.
    \item Evaluated students throughout the quarter.
    \item Led observational sessions and mentored students doing research on the CWU research telescope.
\end{itemize}

\noindent\textbf{Northwest Earth Space Science Pathways (NESSP)} \hfill Ellensburg, Washington\\
\textit{Student Office Assistant} \hfill September 2022 | June 2023
\begin{itemize}
    \item Created robotics curriculum for K-12 student teams.
    \item Hosted week-long overnight summer camps for high school students.
    \item Engaged student teams in engineering challenges mimicking a NASA Mars sample return mission.
    \item Encouraged students to connect a variety of skills in engineering, programming, teamwork, and planning.
\end{itemize}

\noindent\textbf{After-School Robotics Instructor} \hfill North Bend, Washington\\
\textit{Instructor} \hfill 2013 | 2020
\begin{itemize}
    \item Taught K-12 students skills in teamwork, leadership, engineering, programming, and project management.
    \item Mentored competitive FIRST student teams.
    \item Coordinate with other student teachers to provide classes to an entire school district.
\end{itemize}


\section*{LEADERSHIP EXPERIENCE}
\noindent\textbf{CWU Astronomy Club}\\
\textit{President} \hfill September 2023 | Present
\begin{itemize}
    \item Run weekly meetings and set agenda.
    \item Engage with members to gauge club goals and research club interests.
    \item Train members on using and giving planetarium shows.
    \item Host regular public planetarium shows for community members.
    \item Write travel and funding requests for student interests.
    \item Make experiences available to students including the 2023 October eclipse.
    \item Engage with local astronomy research including visits to LIGO.
    \item Engage with local schools to run planetarium shows or bring astronomy equipment.
    \item Participate in advertising events for college and department.
\end{itemize}

\noindent\textbf{CWU Astronomy Club}\\
\textit{Treasurer} \hfill September 2022 | June 2023
\begin{itemize}
    \item Balance budget and write funding requests.
    \item Handle donations and funds raised.
    \item Research school finance guidelines.
\end{itemize}

\noindent\textbf{CWU Society of Physics Students (SPS) Chapter}\\
\textit{Secretary} \hfill September 2022 | June 2023
\begin{itemize}
    \item Keep regular meeting notes on agenda.
    \item Manage club email and schedule.
\end{itemize}

\noindent\textbf{NASA Student Launch Challenge: Wildcat Rocketry Team}\\
\textit{Safety Officer} \hfill September 2022 | June 2023
\begin{itemize}
    \item Handle regular communications with NASA.
    \item Create safety guidelines according to the National Association of Rocketry safety standards.
    \item Enforce safety guidelines during all club activities.
    \item Work with division leaders to ensure safe operation and construction of rocket components.
\end{itemize}

\noindent\textbf{Mount Si High School (MSHS) Robotics Club}\\
\textit{President} \hfill September 2016 | June 2020
\begin{itemize}
    \item Manage outreach activities of three competitive teams.
    \item Work with instructors to get students access to power equipment.
    \item Enforce student safety during meetings and competition.
    \item Host annual, state-wide, competitions through volunteer recruitment.
    \item Coordinate the creation of a school-district wide after school robotics program for K-12 students.
    \item Engage with community members at science fairs and other community stem showcases.
\end{itemize}

\noindent\textbf{MSHS FIRST Robotics Team: Si Borgs}\\
\textit{Team Captain} \hfill September 2016 | June 2020
\begin{itemize}
    \item Work with division leads to coordinate the development of a competitive robot in a timely manner.
    \item Schedule outreach activities and coordinate engagement with other teams.
    \item Encourage gracious professionalism in and out of the club.
\end{itemize}


\section*{ADDITIONAL COURSEWORK}
\begin{tabularx}{1\textwidth}{>{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X }
    % \textbf{Physics Courses}&   \textbf{Math Courses}\\
    \vspace{-0.2in}
    \begin{itemize}
        \item PHYS 292 Exploring Physics Teaching
        \item MATH 314 Probability and Statistics
    \end{itemize}
    & \vspace{-0.2in} \begin{itemize}
        \item MATH 260 Sets and Logic
    \end{itemize}
\end{tabularx}


\section*{HONORS AND DISTINCTIONS}
\textbf{Sigma Pi Sigma Department Nomination} \hfill CWU Department of Physics\\
Nominated for academic excellence, quality of research, volunteer service to the department, and community engagement. \hfill June 2023

\noindent\textbf{FIRST WA Deans List Semi-Finalist}  \hfill Mount Si High School\\
This award recognizes a student who, by the nomination of their team members and mentors, are excellent examples of student leaders who have led their teams and communities to an increased awareness of FIRST and it mission. These students have also achieved personal technical expertise and accomplishment.


% Skills Section
\section*{SKILLS AND CERTIFICATIONS}
\begin{itemize}
    \item Python for scripting, simulations, and data science; MatLab; Mathematica; Git; GitHub; \LaTeX.
    \item Dassault Systemes certificate of Mechanical Engineering in Solidworks.
\end{itemize}

% \section*{REFERENCES}
% \textbf{Prof. John Doe}\\
% \textit{\#Position, \# Department, \# Organization name or University name, \# City, \# Country}\\
% E-mail: \href{mailto:johndoe@XXX.XX}{johndoe@XXX.XX}\\
% Scholar Profiles: \href{related link}{University of XXX - Personal Page} | \href{related link}{Google Scholar} | \href{https://www.linkedin.com/in/xxxxxxxx}{LinkedIn}
\end{document}
