% document style header
\documentclass[a4paper, 12pt]{config/homework}

% import default packages
\usepackage{config/defpackages}

% import custom math commands
\usepackage{config/domath}

% end preamble
\begin{document}

% document title
\noindent
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}X>{\centering\arraybackslash}X>{\centering\arraybackslash}X}
Calvin Sprouse & PHYS474 Homework 2 & Due 2024 Jan 17\\
\midrule
\end{tabularx}

% homework problems begin
\begin{enumerate}
\item Let the wavefunction \(\Psi(x,t)\) be a solution to the time-dependent Schr{\"o}dinger equation when the potential energy is given by \(V(x)\). What is the solution to the Shr{\"o}dinger equation if we now consider a potential of \(V(x) + V_0\) where \(V_0\) is a real positive constant.

\pagebreak
\item A particle is observed in a quantum state described by the wavefunction
\[\Psi(x,t) = A\exp\left(-a\left(\frac{mx^2}{\hbar}+it\right)\right),\]
where \(A\) and \(a\) are real positive constants.
\begin{enumerate}[label = (\alph*)]

\item Normalize \(\Psi \).

% As shown in Griffiths, if \(\Psi \) is normalized at a particular time \(t\) then it is normalized at all times \(t\). We then normalize \(\Psi \) at \(t=0\).
\begin{align*}
1 &= \bint{-\infty}{\infty}{\left|\Psi\right|^2}{x}
\\&= \bint{-\infty}{\infty}{\Psi^* \Psi}{x}
\\&= \bint{-\infty}{\infty}{A\exp\left(-a\left(\frac{mx^2}{\hbar}+it\right)\right) A\exp\left(-a\left(\frac{mx^2}{\hbar}-it\right)\right)}{x}
\\&= A^2 \bint{-\infty}{\infty}{\exp\left(-a\left(\frac{mx^2}{\hbar} + it + \frac{mx^2}{\hbar} - it\right)\right)}{x}
\\&= A^2 \bint{-\infty}{\infty}{\exp\left(-\frac{2am}{\hbar}x^2\right)}{x}
\\&= A^2 \sqrt{\frac{\pi}{\frac{2am}{\hbar}}}.
\end{align*}
Solving for \(A\) we find
\[A = {\left(\frac{2am}{\pi \hbar}\right)}^{1/4}.\]

\pagebreak
\item What is the potential \(V(x)\) that this particle finds itself within?
\(\Psi \) is, by definition, a solution to the Schr{\"o}dinger equation. Thus,
\[i\hbar \diffp{\Psi}{t} = -\frac{\hbar^2}{2m} \diffp[2]{\Psi}{x} + V\Psi.\]
Arranging for the potential energy function \(V\) we get
\[V = \frac{1}{\Psi} \left(i\hbar \diffp{\Psi}{t} + \frac{\hbar^2}{2m} \diffp[2]{\Psi}{x}\right).\]
To make these partial derivatives less threatening, we can begin by writing \(\Psi \) in the form \(\Psi = A \psi(x) \phi(t)\).
The \(A\) component is simply \(A\), as found by normalization.
The exponential term in \(\Psi \) can be separated into an \(x\)-dependent and \(t\)-dependent component;
\[\exp\left(\frac{-amx^2}{\hbar} - ait\right) = \exp\left(-\frac{amx^2}{\hbar}\right)\exp\left(-ait\right),\]
which become \(\psi \) and \(\phi \) respectively. We know
\[\phi(t) = \exp\left(-\frac{iEt}{\hbar}\right) = \exp\left(-ait\right),\]
thus \(a = E / \hbar \). We can now write \(V\) in a more approachable way with ordinary derivatives:
\[V = \frac{1}{A\psi\phi} \left(i\hbar A\psi \diff{\phi}{t} + \frac{\hbar^2}{2m} A\phi \diff[2]{\psi}{x}\right).\]
The ordinary derivatives are
\[\diff{\phi}{t} = \diff{}{t}\left[\exp(-ait)\right] = -ai \exp(-ait) = -ai \phi,\]
and
\[\diff[2]{\psi}{x} =
% \diff[2]{}{x}\left[\exp\left(-\frac{amx^2}{\hbar}\right)\right] =
\diff{\psi}{x}\left[-\frac{2amx}{\hbar} \exp\left(-\frac{amx^2}{\hbar}\right)\right] =
% -\frac{2amx}{\hbar} \psi.
\frac{-2am}{\hbar}\left(1 - \frac{2am}{\hbar}x^2\right)\psi.
\]
Thus the potential energy function \(V\) is given by
\begin{align*}
V &= \frac{1}{A\psi\phi} \left(i\hbar A\psi \diff{\phi}{t} + \frac{\hbar^2}{2m} A\phi \diff[2]{\psi}{x}\right)
\\&= \frac{1}{A\psi\phi} \left(-i\hbar A\psi  \frac{E}{\hbar} i \phi + \frac{\hbar^2}{2m} A\phi \frac{-2Em}{\hbar^2}\left(1 - \frac{2Em}{\hbar^2}x^2\right)\psi\right)
\\&= E - E\left(1 - \frac{2Em}{\hbar^2}x^2\right)
\\&= \frac{2 m E^2 }{\hbar^2} x^2.
\end{align*}

\pagebreak
\item Determine the expectation values \(\braket{x}\), \(\braket{x^2}\), \(\braket{p}\), \(\braket{p^2}\).
\begin{enumerate}[label=\roman*.]

\item The expectation value of \(x\), \(\braket{x}\), is given by
\[\braket{x} = \bint{-\infty}{\infty}{x \left|\Psi\right|^2}{x}\]
which is an odd function evaluated over symmetric limits and therefore
\[\braket{x} = 0.\]

\item The mean square position, \(\braket{x^2}\), is given by
\begin{align*}
\braket{x^2} &= \bint{-\infty}{\infty}{x^2 \left|\Psi\right|^2}{x}
\\&= A^2 \bint{-\infty}{\infty}{x^2 \exp\left(- \frac{2am}{\hbar} x^2\right)}{x}
\\&= \sqrt{\frac{2am}{\pi\hbar}} \frac{\sqrt{\pi}}{2\left(\frac{2am}{\hbar}\right)^{3/2}}
\\&=
\end{align*}

\end{enumerate}

\item Determine the standard deviations for position, \(\sigma_x\), and momentum, \(\sigma_p\).
\item Are your values for \(\sigma_x\) and \(\sigma_p\) consistent with the uncertainty principle?
\end{enumerate}

\pagebreak
\item An electron is trapped in a harmonic quadratic potential. Suppose the expectation value for its position is given by \(\braket{x} = \frac{a}{2}\sin(\omega t)\). Here, \(a\) is a real constant with units of length and \(\omega \) is an angular frequency. What, if anything, can be concluded about the electron's momentum?
\end{enumerate}
\end{document}
